---
title: "Methylation Query"
author: Szi Kay Leung
date: "`r Sys.Date()`"

output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Files 

1) rTg4510_RRBSbetaComplete.Rdata 
2) Phenotype file 
(Transferred from "/gpfs/ts0/projects/Research_Project-191406/isabel/RRBS_new/")

3) DMPsInteractionModel_rTg4510_sig_interaction.csv
(Transferred from "/gpfs/ts0/projects/Research_Project-191406/isabel/RRBS_new/DMPs/rTg4510/DMPsInteractionModel_rTg4510_sig_interaction.csv")

```{r files}
# library
suppressMessages(library("dplyr"))
suppressMessages(library("stringr"))
suppressMessages(library("ggplot2"))
suppressMessages(library(kableExtra))

# load files

# RRBSComplete
dir = "/gpfs/mrc0/projects/Research_Project-MRC148213/sl693/DMP/Updated/"
load(file = paste0(dir, "rTg4510_RRBSbetasComplete.RData"))
RRBS_Phenotype <- read.csv(paste0(dir, "Tg4510_coldata_RRBS.csv"), row.names=1, stringsAsFactors=FALSE)
RRBS_Phenotype$Age_months <- as.factor(RRBS_Phenotype$Age_months)
RRBS_Phenotype$Genotype <- as.factor(RRBS_Phenotype$Genotype)
RRBS_Phenotype$Genotype <- relevel(RRBS_Phenotype$Genotype, "WT")

# Interaction Model 
Annotated = read.csv(paste0(dir, "DMPsInteractionModel_rTg4510_sig_interaction.csv"))
```
## Query
WT and TG methylation across all samples from the RRBS completebetas file appear to not correspond with the mean WT and TG methylation from the interaction model. Which value/direction is correct?

## Using RRBS completebetas file 
Subset for DMP(chr8:120887246) annotated to Irf8, with increased methylation in TG

## {.tabset .tabset-pills}

### All Values
```{r all}
# subset RRBScompletebetas file for DMP 
# Average the methylation value
dat = subset(RRBS_completebetas, rownames(RRBS_completebetas) == "chr8:120887246") %>% reshape2::melt() 
dat %>% kbl() %>% kable_styling("striped", full_width = F) %>% scroll_box(width = "500px", height = "200px")
```

### Mean 

```{r mean}
dat = dat %>% group_by(variable) %>% summarise_at(vars(value), list(Methylation = mean))
dat %>% kbl() %>% kable_styling("striped", full_width = F) %>% scroll_box(height = "200px")
```

### Boxplot 
Merge with phenotype file 

```{r plot}
RRBS_Phenotype = RRBS_Phenotype %>% tibble::rownames_to_column(var = "Sample_ID")
df = merge(dat,RRBS_Phenotype[,c("Sample_ID","Genotype","Age_months")], by.x = "variable", by.y = "Sample_ID") %>% 
  mutate(Age_months = factor(Age_months),Genotype = factor(Genotype, levels = c("WT","TG")), Sample = word(variable,c(1),sep = fixed("_")))

ggplot(df, aes(x = Genotype, y = Methylation)) + geom_boxplot()
```

## Using Interaction Model 
Subset for DMP(chr8:120887246) annotated to Irf8, with decreased methylation in mean TG mice

```{r df}
dat = Annotated[Annotated$betaResultsClean....Position.. == "chr8:120887246",]
dat %>% kbl() %>% kable_styling("striped", full_width = F) %>% scroll_box(width = "100%")
```



