---
title: "DMP in Target Genes"
author: Szi Kay Leung
date: "`r Sys.Date()`"

output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressMessages(library("dplyr"))
suppressMessages(library("DT"))
suppressMessages(library("kableExtra"))

```


```{r cars}
dmp = read.csv("/gpfs/mrc0/projects/Research_Project-MRC148213/sl693/DMP/rTg4510-DMPsPathology_annotated_DiffMeth.csv")
targeted_genes = c("Bin1","Trem2", "Cd33", "Vgf", "Fyn","Mapt","Trpa1", "Picalm", "Sorl1", "Abca7","Snca","Apoe","Abca1","App","Ank1","Clu", "Fus","Ptk2b","Rhbdf2","Tardbp")

```

To check for differentially methylated CpGs from Isabel's RRBS data (Tg4510) in the target genes.

Target Genes: `r targeted_genes`


## {.tabset .tabset-pills}

### Full Table 
No difference whether searching for Gene or Nearest Gene

`r nrow(dmp %>% filter(Gene %in% targeted_genes))` DMP sites annotated to `r 
dmp %>% filter(Gene %in% targeted_genes) %>% group_by(Gene) %>% tally() %>% nrow()` target genes 

```{r}
datatable(dmp %>% filter(Gene %in% targeted_genes) %>% .[,c(20,21,1,3:19,22:27)],
      options = list(scrollX = TRUE))

```

### Number of DMP per target gene

```{r}
kable(dmp %>% filter(Gene %in% targeted_genes) %>% group_by(Gene) %>% tally()) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))
```

